### 2023-07-12 subsections/simulation/poc/poc.tex

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{float}
\usepackage{wrapfig,lipsum}
\usepackage{svg}
\usepackage{mathtools}
\usepackage{tabu}
\usepackage{subcaption}
\usepackage[a4paper, total={6in, 8in}]{geometry}
\usepackage{minted}

\begin{document}

\newcommand{\vecthreeBF}[1]{\vec{\textbf{#1}}}
\newcommand{\vecthree}[1]{\vec{#1}}

\newcommand{\parDeriv}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\parDerivS}[2]{\frac{\partial^2 #1}{\partial #2^2}}
\newcommand{\derivS}[2]{\frac{d^2 #1}{d#2^2}}

\newcommand{\dotProdBF}[2]{\vecthreeBF{#1} \cdot \vecthreeBF{#2}}
\newcommand{\dotProd}[2]{\vecthree{#1} \cdot \vecthree{#2}}

\newcommand{\crossProdBF}[2]{\vecthreeBF{#1} \times \vecthreeBF{#2}}
\newcommand{\crossProd}[2]{\vecthree{#1} \times \vecthree{#2}}

\newcommand{\fromeq}[1]{\textit{equation \ref{eq:#1}}}
\newcommand{\fromeqs}[2]{\textit{equations \ref{eq:#1} and \ref{eq:#2}}}
\newcommand{\fromeqsth}[3]{\textit{equations \ref{eq:#1}, \ref{eq:#2} and \ref{eq:#3}}}

\newcommand{\fromfig}[1]{\textit{figure \ref{fig:#1}}}
\newcommand{\fromfigs}[2]{\textit{figures \ref{fig:#1} and \ref{fig:#2}}}

\newcommand{\fromsec}[1]{\textit{section \ref{sec:#1}}}
\newcommand{\fromsecs}[2]{\textit{sections \ref{sec:#1} and \ref{sec:#2}}}

%----../../..++++.

%%%%%%
% Start of poc.tex
\section{Proof of Concept}
Once the need for a custom built tool for designing and simulating a Rhodotron type accelerator was apparent, a simple 1D proof of concept was implemented.

This proof of concept worked with interacting an electron with the EM field provided, at discrete time steps $dt$. The core sequence of this software is provided below.
Where $\vecthreeBF{r}$ is the position, $\vecthreeBF{v}$ is the velocity of the $e^-$;
\begin{itemize}
    \item Get electron energy $E_{in}$
    \item Get the RF field definitions
    \item Get the magnet design parameters if there is any
    \item Do following until the simulation time has ellapsed
    \begin{enumerate}
        \item Calculate $\vecthreeBF{E}_{||}$($\vecthreeBF{r}$)
        \item Calculate $\vecthreeBF{a}_{E ||}$ using \fromeq{para_and_perp_acc_of_lorentz_force}
        \item Calculate $\vecthreeBF{r}$($t+dt$) using \fromeq{leapfrog_sync_x}
        \item Calculate $\vecthreeBF{v}$($t+dt$) using \fromeq{leapfrog_sync_v}
        \item Calculate new $E_t$
    \end{enumerate}
\end{itemize}

The implemenation of this logic can be observed in \fromfig{POC_core_logic}. 
\begin{figure}[H]
    \begin{minted}[linenos=true, autogobble, frame=lines, framesep=2mm, fontsize=\footnotesize]{c++}
        for (double t=0; t<SimuTime; t+=dT){
            double RelBeta  = v/c;
            double RelGamma = 1.0 / sqrt(1.0-RelBeta*RelBeta);

            double ef=Eradial(r_pos*1000,t,0); // convert position to mm
            double acc=ef*1E6*eQMratio/(RelGamma*RelGamma*RelGamma);

            r_pos = r_pos + v * dT*ns + 1/2*acc*(dT*ns)*(dT*ns);
            v = v + acc*dT*ns;

            RelBeta  = v/c;
            RelGamma = 1.0 / sqrt(1.0-RelBeta*RelBeta);
            Et=RelGamma*E0; 
        }
    \end{minted}
    \caption{Core logic loop of the POC}
    \label{fig:POC_core_logic}
\end{figure}

After simulating the \textit{synchronous electron} mentioned in \fromsec{cavity_of_a_rhodotron} with $\phi_{lag}=15^\circ$ for one pass in THE CAVITY, the results from \textit{POC}, \textit{CST} and \fromeq{W_gain_each_pass_pottier} are compared.
% 40 kW total power
% Emax = 0.96 in old, 2.08 in new version
% R1 = 0.1882
% R2 = 0.753
% f = 107.5
\begin{figure}[H]
    \begin{eqnarray}
        E_{Pottier} &=& 0.565 MeV \nonumber\\
        E_{CST} &=& 0.872 MeV  \label{eq:poc_E_results}\\
        E_{POC} &=& 0.873 MeV \nonumber
    \end{eqnarray}
    \caption*{$P=40$kW, $R_1=0.188$m, $R_2=0.753$m, $f=107.5$MHz}
\end{figure}

As can be observed in \fromeq{poc_E_results}, \textit{POC software} and \textit{CST} produced very similar results. However, we cannot make the same claim regarding \fromeq{W_gain_each_pass_pottier}.
This inconsistency between $E_{Pottier}$ and $E_{CST}$ was noticed in the earlier simulations as well, which supported the idea of using another simulation software to reduce reliance on $CST$.
Using a single software would have made the process error prone. Their results, at the end of the day, depend heavily on the physical and mathematical models they use.

After the promising results obtained from \textit{POC}, the decision was made to develop a more robust simulation software. This software was named \textit{Rhodotron Simulation} for the current stage and started development in October, 2021.
The development and improvement efforts are still ongoing.

During the development of \textit{Rhodotron Simulation} software, several intermediate builds have been implemented and tested. The following sections in this chapter investigates their implementations further.

%%%%%%

-lout opt
-phlag opt
-going 3D
-implement actual magnets instead of relying on lout
-

\end{document}