### 2023-09-19 subsections/simulation/poc/poc.tex

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{float}
\usepackage{wrapfig,lipsum}
\usepackage{svg}
\usepackage{mathtools}
\usepackage{tabu}
\usepackage{subcaption}
\usepackage[a4paper, total={6in, 8in}]{geometry}
\usepackage{minted}

\begin{document}

\newcommand{\vecthreeBF}[1]{\vec{\textbf{#1}}}
\newcommand{\vecthree}[1]{\vec{#1}}

\newcommand{\parDeriv}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\parDerivS}[2]{\frac{\partial^2 #1}{\partial #2^2}}
\newcommand{\derivS}[2]{\frac{d^2 #1}{d#2^2}}

\newcommand{\dotProdBF}[2]{\vecthreeBF{#1} \cdot \vecthreeBF{#2}}
\newcommand{\dotProd}[2]{\vecthree{#1} \cdot \vecthree{#2}}

\newcommand{\crossProdBF}[2]{\vecthreeBF{#1} \times \vecthreeBF{#2}}
\newcommand{\crossProd}[2]{\vecthree{#1} \times \vecthree{#2}}

\newcommand{\fromeq}[1]{\textit{equation \ref{eq:#1}}}
\newcommand{\fromeqs}[2]{\textit{equations \ref{eq:#1} and \ref{eq:#2}}}
\newcommand{\fromeqsth}[3]{\textit{equations \ref{eq:#1}, \ref{eq:#2} and \ref{eq:#3}}}

\newcommand{\fromfig}[1]{\textit{figure \ref{fig:#1}}}
\newcommand{\fromfigs}[2]{\textit{figures \ref{fig:#1} and \ref{fig:#2}}}

\newcommand{\fromsec}[1]{\textit{section \ref{sec:#1}}}
\newcommand{\fromsecs}[2]{\textit{sections \ref{sec:#1} and \ref{sec:#2}}}

%----../../..++++.

%%%%%%
% Start of poc.tex
\section{Proof of Concept}
The proof of concept (POC) worked with interacting an electron with the EM field provided, at discrete time steps $dt$. 
The core sequence of this software is provided below and the implemenation of this logic can be observed in \fromfig{POC_core_logic}. 
Where $\vecthreeBF{r}$ is the position, $\vecthreeBF{v}$ is the velocity of the $e^-$:
\begin{itemize}
    \item Get electron energy $E_{in}$,
    \item Get the RF field definitions,
    \item Get the magnet design parameters if there is any,
    \item Do following until the simulation time has ellapsed,
    \begin{enumerate}
        \item Calculate $\vecthreeBF{E}_{||}$($\vecthreeBF{r}$),
        \item Calculate $\vecthreeBF{a}_{E ||}$ using \fromeq{para_and_perp_acc_of_lorentz_force},
        \item Calculate $\vecthreeBF{r}$($t+dt$) using \fromeq{leapfrog_sync_x},
        \item Calculate $\vecthreeBF{v}$($t+dt$) using \fromeq{leapfrog_sync_v},
        \item Calculate new $E_t$.
    \end{enumerate}
\end{itemize}
\clearpage
\begin{figure}[H]
    \begin{minted}[linenos=true, autogobble, frame=lines, framesep=2mm, fontsize=\footnotesize]{c++}
        for (double t=0; t<SimuTime; t+=dT){
            double RelBeta  = v/c;
            double RelGamma = 1.0 / sqrt(1.0-RelBeta*RelBeta);

            double ef=Eradial(r_pos*1000,t,0); // convert position to mm
            double acc=ef*1E6*eQMratio/(RelGamma*RelGamma*RelGamma);

            r_pos = r_pos + v * dT*ns + 1/2*acc*(dT*ns)*(dT*ns);
            v = v + acc*dT*ns;

            RelBeta  = v/c;
            RelGamma = 1.0 / sqrt(1.0-RelBeta*RelBeta);
            Et=RelGamma*E0; 
        }
    \end{minted}
    \vspace{20pt}
    \caption{Core logic loop of the POC.}
    \label{fig:POC_core_logic}
\end{figure}

After simulating the \textit{synchronous electron} mentioned in \fromsec{cavity_of_a_rhodotron} 
with $\phi_{lag}=15^\circ$, $P_{diss}=40$kW, $R_1=0.188$m, $R_2=0.753$m, $f=107.5$MHz for one pass, 
the results from \textit{POC}, \textit{CST} and \fromeq{W_gain_each_pass_pottier} were found to be
% 40 kW total power
% Emax = 0.96 in old, 2.08 in new version
% R1 = 0.1882
% R2 = 0.753
% f = 107.5
\vspace{-10pt}\begin{eqnarray}
    E_{Pottier} &=& 0.593 MeV  , \\
    E_{CST} &=& 0.521 MeV  ,  \label{eq:poc_E_results}\\
    E_{POC} &=& 0.532 MeV . 
\vspace{-20pt}\end{eqnarray}

As can be observed in \fromeq{poc_E_results}, \textit{POC software} and \textit{CST} produced similar results; $\Delta E_{(POC-CST)}/E_{CST} \approx 2\%$.
while \fromeq{W_gain_each_pass_pottier} was noticably different; $\Delta E_{(Pottier-CST)}/E_{CST} \approx 14\%$.
This inconsistency between \fromeq{W_gain_each_pass_pottier} and  \textit{CST} simulation was noticed in the earlier simulations as well, which supported the idea of using another simulation software to reduce reliance on \textit{CST}.
Since using a single software would have made the process error prone.

After the results obtained from \textit{POC} were found to be promising, a decision was made to develop a more robust simulation software. This software was named \textit{Rhodotron Simulation} for the current stage and started development in October, 2021.
The development and improvement efforts are still ongoing.

During the development of \textit{Rhodotron Simulation} software, several intermediate builds have been implemented and tested. The following sections in this chapter investigates their implementations further.

%%%%%%

-lout opt
-phlag opt
-going 3D
-implement actual magnets instead of relying on lout
-

\end{document}